---
title: "Intro: plot types and basic parameters"
description: >
  Some examples of what psrcplot can do
output: html_vignette
vignette: >
  %\VignetteIndexEntry{Intro: plot types and basic parameters}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## Why **psrcplot**?

Data visualizations can be deceptively complex to get 'just right'. The **psrcplot** library provides a convenient way to create charts with standard PSRC styles, fonts and formatting--without having to specify each of those settings yourself. It uses features and styles through the **ggplot2** package, with optional interactive capabilities via the **plotly** package. The package will also expand and evolve to keep pace with PSRC graphic design standards.

## Visualization menu

Currently, these are the plot types you can create with psrcplot commands:

* Static Bar
* Static Column
* Static Line
* Bubble
* Treemap

[Static plots](2-static-plots.html) such as these are **ggplot** objects and can be edited using [**ggplot2** commands](https://ggplot2.tidyverse.org/reference/gg-add.html).

Additionally, the bar, column and line plot functions each have an [interactive version](3-interactive-plots.html) that adds features such as  tooltips or hovertext (appearing with cursor movement). The result are **plotly** objects, which can be edited using [**plotly** commands](https://plotly.com/r/styling-figures/).

You can find out more about any of the functions by typing ?function_name in the console, e.g.

`?static_bar_chart()`

## Basic parameters

Although each chart type has some nuances, key elements can be generalized across plot types (as those familiar with **ggplot2** will know). The function argument names approximate those in **ggplot2**. The required arguments are:

1. The dataframe/tibble, **t**
2. A discrete variable to separate distinct categories: either **y** (bar plots), **x** (column plots), or **s** (either bubble or treemap plots)
3. A numeric variable plotted as a quantity: either **x** (bar plots), **y** (column plots) or for line plots, both **x** and **y**. 
4. **fill**, which corresponds to the plotted shape: in the case of bubbles or treemap sections, it is numeric and represents size. In bar or column plots, it is a discrete variable; if multiple categories exist they will be listed in the legend. can either be stacked or "dodged" (shown adjacent one another in a group). 

There are also several standard but optional arguments:

* **title** and **subtitle** text, by default located to the top right of the plot
* **source** citation text to appear in smaller font below the plot
* **est** can be specified to adjust the display of the numeric variable listed above (e.g. "currency" will display dollar signs, versus "percent" or "number"); **dec** controls decimal places displayed on the label.
* The **color** argument determines which palette is applied to the fill variable. You can view the standard PSRC alternatives with the function [**`print_psrc_palette()`**](library/prsrplot/html/print_psrc_palette.html).
```{r}
library(psrcplot)

print_psrc_palette()
```

Further optional arguments exist specific to plot type--for example, line width for line plots, or custom axes labels for bar/column plots. 
