---
title: "Customization"
description: >
  Using ggplot2 to tailor your plot - static or interactive
output: html_vignette
vignette: >
  %\VignetteIndexEntry{Customization}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## A good thing only goes so far . . .

Despite our best efforts, **psrcplot** cannot be all things to all people. ¯\\_ (ツ)_/¯ There will always be cases in which the vanilla PSRC standard will need adjustment, either based on the plot content or to fit the publication in which it features. 

Luckily, as **ggplot2** objects, static charts can be extensively tinkered with through [**ggplot2**'s **`+`** syntax](https://ggplot2.tidyverse.org/reference/gg-add.html). Even interactive plots can benefit from this treatment--if done on a static **ggplot2** object (and made into interactive **plotly** objects afterwards).

There are also times where you'll want to handle things like plot titles or captions in the surrounding document rather than as elements of the plot itself. We've found this particularly true of interactive plots, as the alignment options in **plotly** are fairly limited. In such cases, you can simply leave the title and subtitle arguments out.


## Adding or revising static plots with ggplot2

Static psrcplot charts **ggplot** objects and can be edited using [**ggplot2** commands](https://ggplot2.tidyverse.org/reference/gg-add.html). In this example, we create a line chart, then alter the axis scale as well as drop the legend (since that information is in the title).

```{r customize_via_ggplot_example1, message=FALSE}
library(psrcplot)
library(dplyr)

passenger_volume <- enplanement_example_data %>% 
                       filter(loc_id=="SEA") %>% 
                       mutate(data_year=as.character(data_year))

static_plot <- static_line_chart(t=passenger_volume, 
                       x='data_year', y='enplanements', fill='airport_name', 
                       title="Sea-Tac passengers (annual total)",
                       lwidth = 1.5,
                       breaks = seq(2006, 2018, 4))

static_plot
```
```{r customize_via_ggplot_example2, message=FALSE}
static_rev <- static_plot + ggplot2::scale_y_continuous(
                                limits=c(0,25e6), 
                                labels=scales::unit_format(unit="M", scale=1e-6)) + 
                            ggplot2::guides(color="none")

static_rev
```

## Applying PSRC styles to **ggplot2** objects created outside **psrcplot**

For those who create plots directly with ggplot2, you can still apply PSRC formats using the [**`psrc_style()`**](library/prsrplot/html/psrc_style.html) function.


## Making a custom plot interactive

You can use the function [**`psrcplot::make_interactive()`**](library/prsrplot/html/make_interactive.html) to convert a **psrcplot** or **ggplot2** object to an interactive **plotly** object with PSRC styling. Due to constraints in **plotly**, you must currently specify the `source` argument as well. 

```{r make_interactive_example, message=FALSE}
interactive_plot <- static_rev %>% make_interactive()

interactive_plot
```

Since we can't see the future and **plotly** can be a little eccentric, there are no guarantees this command will by itself generate exactly what you want, but it should help. Some further tweaking in **plotly** may be required via the [`plotly::style()` and `::layout()` commands](https://plotly.com/r/creating-and-updating-figures/#Updating%20Figures).