---
title: "Frequently Asked Questions"
description:  FAQs for doing tricky tasks
output: html_vignette
vignette: >
  %\VignetteIndexEntry{Frequently Asked Questions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  rmarkdown.html_vignette.check_title = FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

These examples each use summarized PUMS data, which you can generate yourself using the **psrccensus** package:

```{r example_data_retrieval, message=FALSE}
# library(magrittr)
# library(psrccensus)
# library(dplyr)
# 
# # Retrieve, summarize and filter Census PUMS data
#   z_2019<- get_psrc_pums(span=5,
#                          dyear=2019,
#                          level='h',
#                          vars=c("HRACE", "HINCP"))
#   
#   hhinc19_x_race <- psrc_pums_median(
#                           z_2019, 
#                           stat_var="HINCP",
#                           group_vars="HRACE") %>%
#     filter(HRACE !="Total") %>% mutate(DATE=as.character(DATA_YEAR))

```

## Category display order

This is best done prior to invoking the **psrcplot** function by converting the variable to a Factor datatype with  levels specified in the sequence you'd like them displayed. 

In the case below, the HRACE variable arrives as a Factor datatype, but still with alphabetical default ordering. 

```{r default_ordering, message=FALSE}
library(magrittr)
library(dplyr)
library(psrcplot)
  
hhinc19_x_race <- psrcplot::summary_pums_example_data

# Create chart -- default category ordering   
income.chart.default <- static_bar_chart(
     t=hhinc19_x_race,
     x="HINCP_median", y="HRACE",
     fill="DATE",
     moe='HINCP_median_moe',
     est='currency',
     title="Median Household Income by Race - default Race order")

income.chart.default

```

Factor levels can be reversed using the `fct_rev()` function from the **forcats** package.

```{r reverse_ordering, message=FALSE}
hhinc19_rev <- hhinc19_x_race %>% 
  mutate(HRACE_REV=forcats::fct_rev(as.factor(HRACE)))

income.chart.reverse <- static_bar_chart(
     t=hhinc19_rev, 
     x="HINCP_median", y="HRACE_REV",
     f="DATE",
     moe='HINCP_median_moe',
     est='currency',
     title="Median Household Income by Race - reverse default")

income.chart.reverse

```

### Sort by value

You may wish to order the categories by the data variable instead. This uses another **forcats** function, `fct_reorder()` (more complex multi-variable ordering expressions are also possible using this function).

```{r data_ordering, message=FALSE}
 hhinc19_low_high <- hhinc19_x_race %>% 
  mutate(HRACE_low_high=forcats::fct_reorder(HRACE, -HINCP_median))

income.chart.low_high <- static_bar_chart(
      t=hhinc19_low_high, 
      x="HINCP_median", y="HRACE_low_high",
      fill="DATE",
      moe='HINCP_median_moe',
      color='psrc_dark',
      est='currency',
      title="Median Household Income by Race - median income order")

income.chart.low_high

```

### Sort by a user-specified order

Sometimes there is an logic that can't be captured by alphabetical or numeric rank--for example, setting remainder or combination categories last. To do that, you'll need to order the factor levels manually. Be careful that you type the exact factor names, or they will be omitted (ugh).

```{r custom_ordering, message=FALSE}
hhinc19_custom <- hhinc19_x_race %>%
    mutate(HRACE_custom=factor(HRACE, 
       levels = c("American Indian or Alaskan Native Alone", 
                  "Asian alone",
                  "Black or African American alone",
                  "Hispanic or Latino",
                  "Native Hawaiian and Other Pacific Islander alone",
                  "White alone",
                  "Some Other Race alone",
                  "Two or More Races", 
                  "Multiple Races")))

income.chart.custom <- static_bar_chart(
      t=hhinc19_custom, 
      x="HINCP_median", 
      y="HRACE_custom",
      fill="DATE",
      moe='HINCP_median_moe',
      color='psrc_dark',
      est='currency',
      title= "Regional Household Median Income by Race - custom order")

income.chart.custom

```


## Further text wrapping for axis labels

By default, **psrcplot** wraps lengthy column labels and leaves horizontal bar labels alone. If you want text wrapping, however, you can add it wherever it is necessary using **ggplot2** and **stringr**. You may need to tinker with the wrap length to achieve what you want.

```{r wrap_labels, message=FALSE}

income.bar <- static_bar_chart(
      t=hhinc19_custom, 
      x="HRACE", 
      y="HINCP_median",
      fill="DATE",
      moe='HINCP_median_moe',
      est='currency')

income.bar

income.column.wrapped <- income.column + 
  ggplot2::scale_y_discrete(labels=function(x) stringr::str_wrap(x, width=18))

income.column.wrapped

```

By inserting the expression in the label rather than altering the object beforehand, this leaves the category variable in the data alone (`str_wrap` would otherwise convert the type to character and lose factor level ordering, etc).
