---
title: "Install the Poppins font"
description: >
  Permanently install the default PSRC font
output: html_vignette
vignette: >
  %\VignetteIndexEntry{Install the Poppins font}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The **psrcplot** package uses PSRC's default font, Poppins. 

The package includes a way for R to access it temporarily; this leaves some stuff behind in your `user\AppData\local\Temp` directory. Since the agency uses Poppins so extensively, it makes sense for staff to install it permanently--however, you'll need to enable all users access to it if R is to utilize that version rather than go back to its own temporary install.

Here are the steps to follow:

  1. Download the font family (a .zipfile archive) from https://fonts.google.com/specimen/Poppins.
  2. Extract/unzip the folder, select all the included font files, and **right-click**
  3. Select ***"Install for all users"*** - without this, R will not see it.
  4. Reboot

The font should be installed in the `C:\Windows\Fonts` directory. 
To verify it is visible to R, confim the following returns `TRUE`:

```{r mode_data, message=FALSE}

"Poppins" %in% sysfonts::font_files()$family
```

