---
title: "Make a static plot"
description: >
  Use the primary functions of psrcplot to create a static bar plot.
output: html_vignette
vignette: >
  %\VignetteIndexEntry{Make a static plot}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

As mentioned in the introduction, psrcplot has static versions of bar, column, line, bubble, and treemap plots. Here we illustrate using a simple bar plot, with American Community Survey (ACS) data from the Census Bureau.

# Prepare & examine the underlying data

The example data measure mode share by race for the PSRC region:

```{r mode_data, message=FALSE}
library(psrcplot)
library(dplyr)
library(ggplot2)

mode_shares <- mode_share_example_data %>% 
 filter(Category=="Mode to Work by Race" & Geography=="Region" & Race=="Total") %>%
 select(-c(Category, Geography, Race)) %>% mutate(Year = as.character(Year))

head(mode_shares)
```

## Using psrccensus `static_??_chart()` functions

To create a bar plot, call the function `static_bar_chart()`, specifying **`t`** as the underlying table. Note that for bar charts the **x** variable should be numeric (as quantities are represented on the x axis) and the **y** variable should be discrete/categorical.

```{r static_plot_modes, message=FALSE}
modes_chart <- static_bar_chart(
                    t=mode_shares, y="Mode",x="share", fill="Year",
                    title="Mode Share to Work",
                    alt="Chart of Work Mode Shares",
                    source=paste("Source: ACS 5-Year Estimates, table B3002",
                                 "for King, Kitsap, Pierce and Snohomish counties.",
                                 sep = "\n"),
                    color=psrcplot::psrc_colors$pgnobgy_5)

modes_chart
```


## Exporting a static chart

To save a static chart programmatically, use the **[`ggplot2::ggsave()`](https://ggplot2.tidyverse.org/reference/ggsave.html)** function. You can also use the Plots->Export menu or right-click -> "Save as" in RStudio.
```{r}
ggsave(filename='modes_bar_chart.png', plot=modes_chart, device='png')

```
